\mythesistitleNOTSYNPOSIS

### 2012-01-11 src/020chap2.tex
made modes in section for PQR trees

Thus, effectively, there exists a bijection between set of matrices
with \COP and the set of \PQtrees (accurately speaking, each matrix
with \COP bijectively maps to an equivalence class of \PQtrees due to
properties (\ref{def::nodep}) and (\ref{def::nodeq})). See
Figure~\ref{fig:pqtree} for an example of \PQtree.

@@

This algorithm is notoriously difficult to program. In the procedure
to change the \PQtree at each iteration, nodes are considered from
leaves to tree. At each node considered, it uses one of nine templates
to determine how the tree must be altered in the vicinity of this
node. Recognition of this template poses a difficult challenge in
terms of implementing it. Each template is actually a representative
of a larger class of similar templates, which must be dealt with
explicitly by a program\cite{mcc04}.

@@

@@

%\clearpage      % Print pending floats
%\section{Certifying Algorithms}
\subsection{Generalized \PQtree\ --  COP for set systems}
\label{sec:surveycertalgo}

Section~\ref{sec:motive} mentions how a binary matrix naturally maps
to a system of sets.  A set can be constructed for each column of
matrix with its elements being those row indices at which the column
has \un s. Thus the collection of sets corresponding each column of
the matrix forms a set system with universe as the set of all row
indices of the matrix.  This simple construction is formally described
in Definition~\ref{def:matrixsetsystem}.

\begin{definition}%
  \label{def:matrixsetsystem}%
  [\emph{Set system of a binary matrix}.] Let $M$ be a binary matrix
  of order $n \times m$ % , $r_i, i \in [n]$ be the rows in $M$
  and $\{c_i \mid i \in [m]\}$ be the columns in $M$. A set system
  $\cF_{M} = \{S_i \mid S_i \subseteq [n], i \in [m]\}$ is defined
  such that for every column $c_i$ of $M$, set $S_i = \{j \mid m_{ji}
  = \un \}$. The collection $\cF_M$ is the set system of matrix $M$.
\end{definition}

Thus the idea of \cop naturally maps to set sytems as well\endnote{As
  seen in Section~\ref{sec:courseschedule}} as
Definition~\ref{def:copforsets} states. This is equivalent to \COP for
matrices seen in Definition~\ref{def:copmatrix}.

\begin{definition}%
  \label{def:copforsets}%
  [\emph{\Cop for set systems% \cite{mcc04}
  }.] A set system $\cF$ with universe $U$ has the \cop if it is
  possible to assign a linear order to $U$ where each set $S \in \cF$
  is an interval in the linear order.\tnote[important]{NOT formal
    enough.}
\end{definition}

\tnote{What does n89 say? download paper from cornell website.}
Generalized \PQtree is a data structure defined in \cite{mcc04} for
arbitrary set systems. If the set system has \COP, the generalized
\PQtree is identical to its \PQtree.

An important observation made by \cite{mm96} is that if $\cF$ is a set
system with \COP then, aside from every set in $\cF$ being consecutive
(after applying the \COP order) the following sets must also be
consecutive for any $A, B \in \cF$ 
\begin{enumerate}
  \item The intersection $A \cap B$ 
  \item The union $A \cup B$ if $A \cap B \ne \emptyset$
  \item The relative complements $A \setminus B$ and $B \setminus A$
    if $B \nsubseteq A$ and  $A \nsubseteq B$ respectively.
  \item Additionally the sets $U$, singleton sets and the empty set $\emptyset$
    (collectively called {\em trivial sets} in \cite{mm96}) are consecutive in all orderings
    of the universe. $\emptyset$ is considered consecutive by convention.
\end{enumerate}

% \section{COP in Relational Database Model}
% \label{sec:surveyrdbm}
% \tnote{Expand on sec:apprdbm} \tnote{(set systems theme)}
\section{Optimization problems in COP}
\label{sec:surveycopopt}
\tnote{Expand on ref:sec:optcop}

>>>>

See Figure~\ref{fig:pqtree} for an example of \PQtree. It may be noted
that there is no way an empty set of \COP orderings can be represented
in this data structure. For this reason, \PQtree is undefined for
matrices that do not have \COP.  Thus effectively, there exists a
bijection between set of matrices with \COP and the set of \PQtrees
(accurately speaking, each matrix with \COP bijectively maps to an
equivalence class of \PQtrees resulting from properties (\ref{def::nodep}) and
(\ref{def::nodeq})).

@@

Judging from notes in literature, this algorithm is apparently
notoriously difficult to program. In the procedure to modify the
\PQtree at each iteration, nodes are considered from leaves to
tree. At each node considered, it uses one of nine templates to
determine how the tree must be altered in the vicinity of this
node. Recognition of this template poses a difficult challenge in
terms of implementing it. Each template is actually a representative
of a larger class of similar templates, which must be dealt with
explicitly by a program\cite{mcc04}.

@@

%\section{Certifying Algorithms}
\subsection{\PQRtree\ --  COP for set systems}
\label{sec:surveycertalgo}

Section~\ref{sec:motive} mentions how a binary matrix naturally maps
to a system of sets.  A set can be constructed for each column of
matrix with its elements being those row indices at which the column
has \un s. Thus the collection of sets corresponding each column of
the matrix forms a set system with universe as the set of all row
indices of the matrix.  This simple construction is formally described
in Definition~\ref{def:matrixsetsystem} along with the idea of \cop
for set systems\endnote{As seen in Section~\ref{sec:courseschedule}}.

\begin{definition}%
  \label{def:matrixsetsystem}%
  % [\emph{Set system of a binary matrix}.]
  [\emph{\Cop for set systems}.]  Let $M$ be a binary matrix of order
  $n \times m$ % , $r_i, i \in [n]$ be the rows in $M$
  and $\{c_i \mid i \in [m]\}$ be the columns in $M$. A set system
  $\cF_{M} = \{S_i \mid S_i \subseteq [n], i \in [m]\}$ is defined
  such that for every column $c_i$ of $M$, set $S_i = \{j \mid m_{ji}
  = \un \}$. The collection $\cF_M$ is the {\em set system of binary
    matrix} $M$.\\
  A set system $\cF$ from universe $U = \{1, 2, \ldots, n\}$ has the
  \cop if it is possible to assign a linear order $\lambda$ to $U$
  where each set $S \in \cF$ appears as a consecutive subsequence of
  $\lambda$ (an interval) in the linear order.
\end{definition}

It is easy to see the equivalence of this definition to \COP for
matrices in Definition~\ref{def:copmatrix}.

Generalized \PQtree is a data structure defined in \cite{n89, mcc04} for
arbitrary set systems. If the set system has \COP, the generalized
\PQtree is identical to its \PQtree. We will use the notation from
\cite{n89} and call generalized \PQtrees as \gPQtrees.

A \gPQtree of a given set system $\cF$ from universe $U$ represents
all the sets from $\power{U}$ that have trivial intersections with
every set in $\cF$.

Two sets $A$ and $B$ are said to have a trivial intersection if $A
\cap B$ is one of the following -- $\emptyset$, $A$ or $B$. In other
words, $A$ and $B$ are either disjoint or one is the subset of the
other. {\em Trivial sets} are sets that have trivial intersections
with any set in $\power{U}$. These are namely, $U$, singleton sets in
$\power{U}$ and $\emptyset$\cite{n89, mm96}.

An important observation made by \cite{mm96} is that if $\cF$ is a set
system with \COP then, aside from every set in $\cF$ being consecutive
(after applying the \COP order) the following sets must also be
consecutive for any $A, B \in \cF$.
\begin{enumerate}
  \item The intersection $A \cap B$ 
  \item The union $A \cup B$ if $A \cap B \ne \emptyset$
  \item The relative complements $A \setminus B$ and $B \setminus A$
    if $B \nsubseteq A$ and $A \nsubseteq B$ respectively.
\end{enumerate}

Additionally the {\em trivial sets} of universe $U$ denoted by
$\cT(U)$, are namely -- $U$, singleton sets and the empty set
$\emptyset$. Clearly, $\cT(U)$ are consecutive
in all orderings of $U$. $\emptyset$ is considered consecutive by
convention.

Overlap components are computable in linear time \cite{mm95,wlh92}.

@@

%\clearpage      % Print pending floats
% \section{COP in Relational Database Model}
% \label{sec:surveyrdbm}
% \tnote{Expand on sec:apprdbm} \tnote{(set systems theme)}
\section{Optimization problems in COP}
\label{sec:surveycopopt}
\tnote{Expand on ref:sec:optcop}
